A graduate introduction to classic McEliece

### 2024-11-08 code-based-crypto/talk/main.tex
more references

>>>>

\section{Preliminaries}
% TODO: degree of polynomial
% TODO: Hamming weight
% TODO: Bernoulli's rule

\section{Understanding binary Goppa code decoding}
\subsection{Polynomial interpolation}

\begin{theorem}[Lagrange Interpolation]\label{thm:lagrange-interpolation}
    Let $K$ be a finite field, $\alpha_1, \alpha_2, \ldots, \alpha_n \in K$ be $n$ distinct field elements, and $r_1, r_2, \ldots, r_n \in K$ be $n$ (possibly non-distinct) elements. There exists a unique degree-$(n-1)$ polynomial $f$ such that $f(\alpha_i) = r_i$:

    \begin{equation*}
        f = \sum_{i=1}^n\left(
            r_i \prod_{j \neq i}\frac{x - \alpha_j}{\alpha_i - \alpha_j}
        \right)
    \end{equation*}
\end{theorem}

It's easy to check that the interpolation is correct, and that the degree of $f$ is less than $n$. If there exists another $g \in K[x]$ with $\deg(g) < n$ that interpolates $(\alpha_i, r_i)_{i=1}^n$, then $f - g$ is a polynomial with $n$ distinct roots. However, $\deg(f-g)$ is less than $n$ since $\deg(f), \deg(g) < n$. Therefore, it must be that $f-g = 0$, which means that $f = g$, thus proving uniqueness.

\subsection{Shamir's secret sharing}
Lagrange interpolation can be used to build a secret sharing scheme \cite{shamir1979share}.

% TODO: shamir secret sharing

\subsection{Polynomial approximant}
% TODO: need intuition?

\begin{theorem}[Best approximant theorem]\label{thm:best-approximant-theorem}
    Let $K$ be some finite field and $n, t$ be non-negative integers such that $2t < n$. Given polynomials $A, B \in K[x]$ such that $\deg(B) < \deg(A)$, then there exists unique pair of polynomials $a, b \in K[x]$ such that $\deg(a) \leq t$, $\deg(b) < \deg(a)$, $\gcd(a, b) = 1$, and $\deg(aB - bA) < n - t$. If $c, d \in K[x]$ is such that $\deg(c) \leq t$ and $\deg(cB - dA) < n - t$, then $(c, d) = \lambda (a, b)$ for some $\lambda \in K[x]$.
\end{theorem}

% TODO: proof

\subsection{Interpolation with error}
In this section we introduce an algorithm that can recover a degree $(n - 2t)$ polynomial given $n$ points with up to $t$ errors.

Let $K$ be some finite field and $n, t$ be non-negative integers such that $2t < n$. Let $f \in K[x]$ be a polynomial with degree $\deg(f) < n - 2t$. Let $\alpha_1, \alpha_2, \ldots, \alpha_n$ be $n$ distinct elements. Denote $\mathbf{c} = (f(\alpha_1), f(\alpha_2), \ldots, f(\alpha_n)) \in K^n$. If $\mathbf{r} \in K^n$ is such that $\mathop{wt}(\mathbf{r} - \mathbf{c}) \leq t$, then the following procedure can recover $f$ using $\alpha_1, \alpha_2, \ldots, \alpha_n$ and $\mathbf{r}$: \begin{enumerate}
    \item Let $A = \prod_{i=1}^n (x - \alpha_i)$
    \item Let $B$ interpolate $(\alpha_i, r_i)$
    \item Compute degree-$t$ approximant $(a, b)$ of $(A, B)$. The error can be corrected if and only if $a \mid A$
    \item Compute $f = B - bA / a$. Furthermore, $(B - f)(\alpha_i) \neq 0$ if and only if $a(\alpha_i) = 0$.
\end{enumerate}

\begin{proof}
    % TODO: proof
\end{proof}

\subsection{Goppa decoding}

\subsection{Parity check matrix}
